---
title: "Storm Data Analysis"
author: "Chandrasekar Ganesan"
date: "March 11, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction

##Background
Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

##Data References

The data for this assignment come in the form of a comma-separated-value file compressed via the bzip2 algorithm to reduce its size. You can download the file from the course web site:

* [Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)

There is also some documentation of the database available. Here you will find how some of the variables are constructed/defined.

* National Weather Service [Storm Data Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)
* National Climatic Data Center Storm Events [FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)

The events in the database start in the year 1950 and end in November 2011. In the earlier years of the database there are generally fewer events recorded, most likely due to a lack of good records. More recent years should be considered more complete.

##Key Questions

* Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

* Across the United States, which types of events have the greatest economic consequences?

#Data Analysis 

## Get Data

```{r cache=TRUE}

tf <- tempfile()
file_url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(file_url, destfile = tf)

stormdata <- read.csv(tf, stringsAsFactors = FALSE)

```

## Analyze Data

```{r}
head(stormdata)
summary(stormdata)

# Read unique type of data
unique_eventtypes <- unique(stormdata$EVTYPE)

# Read the Injuries data
head(stormdata$INJURIES)
summary(stormdata$INJURIES)
head(stormdata$FATALITIES)
summary(stormdata$FATALITIES)

```

## Analysis

### Question 1: Which event is most harmful to population health?

```{r}
library(ggplot2)

casualities <- with(stormdata, aggregate(INJURIES + FATALITIES ~ EVTYPE, data=stormdata, FUN = "sum"))
names(casualities) <- c("EVTYPE","TotalCasualities")

#sort descending
casualities <- casualities[order(-casualities$TotalCasualities),]
top5 <- head(casualities, 5)

g <- ggplot(top5, aes(EVTYPE, TotalCasualities))
g <- g + geom_bar(stat = "identity", aes(colour=EVTYPE, fill=EVTYPE), show.legend = FALSE)
g <- g + scale_x_discrete(limits = top5$EVTYPE[order(-top5$TotalCasualities)])
g <- g + ggtitle("Event causing most harm to the population health")
g <- g + xlab("Event Type")
g + ylab("Total Casualities")

```

> As it can be observed, the most harmful event is ** TORNADO **

### Question 2: Which event have the greatest economic consequences? 

```{r}
# Get the subset of strom data that deals with costs

stormsubset <- subset(stormdata, (stormdata$CROPDMGEXP == "M" | stormdata$CROPDMGEXP == "K" | stormdata$CROPDMGEXP == "B") | (stormdata$PROPDMGEXP =="M" | stormdata$PROPDMGEXP =="K" | stormdata$PROPDMGEXP =="B") )


# Factor the crop damage data based on codes

for(i in 1:length(stormsubset$CROPDMGEXP)) {
                ifelse(stormsubset$CROPDMGEXP[i] == "M", stormsubset$CROPDMG[i] <- stormsubset$CROPDMG[i] * 1000000, 
                ifelse(stormsubset$CROPDMGEXP[i] == "K", stormsubset$CROPDMG[i] <- stormsubset$CROPDMG[i] * 1000,
                ifelse(stormsubset$CROPDMGEXP[i] == "B", stormsubset$CROPDMG[i] <- stormsubset$CROPDMG[i] * 1000000000, stormsubset$CROPDMG[i] <- stormsubset$CROPDMG[i] * 1)))
}
summary(stormsubset$CROPDMGEXP)

# Factor the property damage data based on codes
for(i in 1:length(stormsubset$PROPDMGEXP)) {
                ifelse(stormsubset$PROPDMGEXP[i] == "M", stormsubset$PROPDMG[i] <- stormsubset$PROPDMG[i] * 1000000,
                ifelse(stormsubset$PROPDMGEXP[i] == "K", stormsubset$PROPDMG[i] <- stormsubset$PROPDMG[i] * 1000,
                ifelse(stormsubset$PROPDMGEXP[i] == "B", stormsubset$PROPDMG[i] <- stormsubset$PROPDMG[i] * 1000000000, stormsubset$PROPDMG[i] <- stormsubset$PROPDMG[i] * 1)))
}
summary(stormsubset$PROPDMGEXP)
```



```{r}
library(ggplot2)
library(scales)

economic <- with(stormsubset, aggregate(CROPDMG + PROPDMG ~ EVTYPE, data=stormsubset, FUN=sum))
names(economic) <- c("Event.Type", "Economic.Damage")

top5_eco <- head(economic[order(-economic$Economic.Damage),], 5)
g
g <- ggplot(top5_eco, aes(Event.Type, Economic.Damage))
g <- g + geom_bar(stat = "identity", aes(colour=Event.Type, fill=Event.Type), show.legend = FALSE)
g <- g + scale_x_discrete(limits = top5_eco$Event.Type[order(-top5_eco$Economic.Damage)])
g <- g + ggtitle("Event causing most economic damage")
g <- g + xlab("Event Type")
g + ylab("Economic Damage") + scale_y_continuous(labels = comma)

```

